Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-13 concl.tex
notes to liz

No forecast strategy will be effective if a time series has no
predictive structure.  In practice, this almost-tautological statement
poses real problems; without knowing anything about the generating
process, it is difficult to determine how much predictive structure is
present in a given time series.  And even if predictive structure
exists, a given forecast method may not work, simply because it cannot
capture and/or exploit the structure that is present (e.g., a linear
model of a nonlinear process).  This too can be a serious practical
challenge.  If a forecast model is not producing good results, a
practitioner needs to know why: is the reason that the time series
contains no predictive structure---i.e., that no model will work---or
is the model that s/he is using simply not good enough?

@@

An important practical corollary to this empirical correlation of
predictability and WPE is a practical strategy for assessing
appropriateness of forecast methods.  If the forecast produced by a
particular method is poor but the time series contains a significant
amount of predictive structure, one can reasonably conclude that that
method is inadequate to the task and that one should seek another
method.  The nonlinear LMA method, for instance, performs better in
most cases because it is more general.  The \naive ~method, which
simply predicts the mean, works better in noisy signals because it
effects a filtering operation.  The dashed line in
Figures~\ref{fig:wpe_vs_mase_best} and~\ref{fig:wpe_vs_mase_all} is a
useful empirical guideline for knowing when a model is not
well-matched to the task at hand: a $MASE$ score that is more than
{\color{red}[[what's the equation for that line, Ryan?]]} times the
WPE of the corresponding signal is an indication that that time series
has more predictive structure than that forecast model can capture and
exploit.

@@

Nonstationarity is a serious challenge in any time-series modeling
problem.  Indeed, one of the first uses of permutation entropy was
regime-shift detection for the purposes of recognizing seizure onset
in brainwave data~\cite{cao2004det}.  Recall that the signal in
Figure~\ref{fig:svd_ts_colored} was especially useful for the study in
this paper because it contained a number of different regimes.  We
segmented this signal visually, but one could imagine using
permutation entropy to do so instead.  Automating regime-shift
detection would be an important step towards a fully adaptive modeling
strategy, where old models are discarded and new ones are rebuilt
whenever shifts are detected.  WPE would be particularly powerful in
this scenario, as its value can suggest what kind of model might work
well in each new regime.

>>>>

No forecast strategy will be effective if a time series has no
predictive structure.{\color{red} no forecast strategy that uses **structure** will be effective, but naive forecast may do just fine. Maybe something like ..."Forecast strategies which depend on predictive structure (e.g., LMA or ARIMA) are rendered ineffective when signal complexity outweighs information redundancy."}  In practice, this almost-tautological statement
poses real problems; without knowing anything about the generating
process, it is difficult to determine how much predictive structure is
present in a given time series.  And even if predictive structure
exists, a given forecast method may not work, simply because it cannot
capture and/or exploit the structure that is present (e.g., a linear
model of a nonlinear process).  This too can be a serious practical
challenge.  If a forecast model is not producing good results, a
practitioner needs to know why: is the reason that the time series
contains no predictive structure---i.e., that no model will work---or
is the model that s/he is using simply not good enough?

@@

An important practical corollary to this empirical correlation of
predictability and WPE is a practical strategy for assessing
appropriateness of forecast methods.  If the forecast produced by a
particular method is poor but the time series contains a significant
amount of predictive structure, one can reasonably conclude that that
method is inadequate to the task and that one should seek another
method.  The nonlinear LMA method, for instance, performs better in
most cases because it is more general.  The \naive ~method, which
simply predicts the mean, works better in noisy signals because it
effects a filtering operation.{\color{red} might be helpful to say \naive wins on \svdone with the exception to the 5 outliers and that random walk beats everyone on \gcc, this shows the two simple predictions beat everyone for the two highly complex signals but might be too late in the paper to say this, but might be able to say this to reinforce it}   The dashed line in
Figures~\ref{fig:wpe_vs_mase_best} and~\ref{fig:wpe_vs_mase_all} is a
useful empirical guideline for knowing when a model is not
well-matched to the task at hand: a $MASE$ score that is more than
{\color{red}[[what's the equation for that line, Ryan?]]} times the
WPE of the corresponding signal is an indication that that time series
has more predictive structure than that forecast model can capture and
exploit.

@@

Nonstationarity is a serious challenge in any time-series modeling
problem.  Indeed, one of the first uses of permutation entropy was
regime-shift detection for the purposes of recognizing seizure onset
in brainwave data~\cite{cao2004det}.  Recall that the signal in
Figure~\ref{fig:svd-ts-colored} was especially useful for the study in
this paper because it contained a number of different regimes.  We
segmented this signal visually, but one could imagine using
permutation entropy to do so instead.  Automating regime-shift
detection would be an important step towards a fully adaptive modeling
strategy, where old models are discarded and new ones are rebuilt
whenever shifts are detected.  WPE would be particularly powerful in
this scenario, as its value can suggest what kind of model might work
well in each new regime.
